On the use of decision trees to identify the local extrema in parallel global optimization algorithms

### 2023-01-31 2023/PAVT/ParallelDecisionTree.tex
Поправлен список литературы.

The main idea of the approach to constructing more efficient nonuniform grid is that its points are  computed sequentially while the objective function is considered as a realization of some random process. The decision rules of the grid building algorithm are constructed in such a way that the next  grid point corresponds to the global minimum point of the mathematical expectation of the function  values. This point is stored in the list of known values, and the iterations are repeated until the stop criteria is satisfied: either the distance between the trial points becomes less than a  predefined value or a preset maximum number of iterations is achieved \cite{fio_bib10}.

@@

The substantiation of this method of organization of computations see in \cite{Strongin2013}. The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \cite{Barkalov2002,Gergel1997}.

@@

The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods. Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{Himmelblau75}.

@@

When implementing the algorithm for finding the attraction areas of the local  minima, we used the algorithms from OpenCV library to construct decision tree. OpenCV is an open source library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms. More details on the decision tree can be found  in \cite{Opencv}.

@@

Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented. The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{fio_bib19}) of dimensionalities 2, 3, 4, and 5. The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.  In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of time speedup relative to the sequential run are compared. The parallelization was performed using MPI technology, the run was performed on 8 processes.

>>>>

The main idea of the approach to constructing more efficient nonuniform grid is that its points are  computed sequentially while the objective function is considered as a realization of some random process. The decision rules of the grid building algorithm are constructed in such a way that the next  grid point corresponds to the global minimum point of the mathematical expectation of the function  values. This point is stored in the list of known values, and the iterations are repeated until the stop criteria is satisfied: either the distance between the trial points becomes less than a  predefined value or a preset maximum number of iterations is achieved.

@@

The substantiation of this method of organization of computations see in \cite{Strongin2000,Barkalov2016}. The  modifications taking into account the presence of inequality constraints as well as the information on the derivative of the objective function are presented in \cite{Barkalov2002,Gergel1997}.

@@

The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods. Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{Himmelblau72}.

@@

When implementing the algorithm for finding the attraction areas of the local  minima, we used the algorithms from OpenCV library to construct decision tree. OpenCV is an open source library of the algorithms of computer vision, image processing, and  general-purpose numerical algorithms. More details on the decision tree can be found  in \cite{Brahmbhatt2013}.

@@

Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented. The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{Sergeyev2006}) of dimensionalities 2, 3, 4, and 5. The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.  In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of time speedup relative to the sequential run are compared. The parallelization was performed using MPI technology, the run was performed on 8 processes.
